Diffusion-limited mixing by incompressible flows

### 2017-10-04 nonlinearity/manuscript.tex
made charlies revisions to subsection Linfinity speed limited flow constraint analysis

\begin{abstract}
Incompressible flows can be effective mixers by appropriately advecting a passive tracer to produce small filamentation length scales. The Batchelor length scale \cite{Batchelor1959a} was proposed as the lower limit of length scales achievable in turbulent flows. Here, we provide numerical evidence that this limitation may be even more general, applying as well to all incompressible flows under mild physical constraints such as bounded enstrophy or energy. We consider local-in-time flow optimization under these flow intensity constraints with the objective of maximizing mixing rate performance. We observe that, for enstrophy-bounded optimal flows, the strength of diffusion has no impact on the long-term mixing rate performance. For energy-constrained optimal flows, however an increase in the strength of diffusion decreases the mixing rate. We provide analytical lower bounds on mixing rates and length scales achievable under related constraints (pointwise bounded speed and rate-of-strain) by extending the work of \cite{JFM2011} and \cite{Chi-Cheu1996}.

\end{abstract}
%
\pacs{47.85.lk, 47.85.L-, 42.10, 47.57.eb}
\ams{76R50, 37A25, 76F25, 76D55}
%
% Uncomment for keywords
\vspace{2pc}
\noindent{\it Keywords}: Mixing, incompressible flow, diffusion processes, flow control and optimization, Batchelor scale

@@

Mixing a passive tracer concentration via an incompressible fluid flow is important to many areas of science and engineering. A question of interest across these domains is ``How can one stir to enhance the rate of mixing?'' We explore this question and study the impact of molecular diffusion. We address this topic by considering an incompressible flow with mild physical constraints in a periodic box with side length $L$ in $d$ dimensions. We consider a mean-zero tracer concentration field $\theta$ that evolves according to the advection-diffusion equation,
\begin{equation}
	\label{eq:PDE_advection}
	\ppt{\theta}+\mathbf{u}\cdot \nabla \theta=\kappa \lap\theta,
\end{equation}
with initial data $\theta(\mathbf{x},0)=\theta_{0}(\mathbf{x})$, where $\kappa$ is the molecular diffusion coefficient and $\mathbf{u}(\mathbf{x},t)$ is a incompressible ($\nabla\,\cdot\, \vec{u}=0$) flow field. The flow is constrained by enstrophy $\ltwo{\nabla\u} = \Gamma L^{d/2}$ or energy $\ltwo{\u} = UL^{d/2}$ where $\Gamma$ is the root mean square rate-of-strain and $U$ is the root mean square speed. We primarily use the $H^{-1}$ norm or mix-norm throughout to measure homogenization,

\begin{equation}
\hmone{\theta}=\sqrt{\sint{ |\nabla^{-1} \theta( \vec{x},t)|^2}}=\sqrt{ \sum_{\vec{k}\neq \vec{0}} L^d \frac{|\hat{\theta}_{\vec{k}}(t)|^{2}}{|\vec{k}|^2}}
\end{equation}
where $\nabla^{-1}=\nabla \Delta^{-1}$, the operator $\Delta^{-1}$ acting on  a function $\rho$ returns the solution $\phi$ of the Poisson equation $ \Delta \phi = \rho $, and $\hat{\theta}_{\vec{k}}(t) =  \frac{1}{L^{d}}\sint{\theta(\vec{x},t)e^{-i\vec{k}\cdot\vec{x}}}$.  Lower values of the  $H^{-1}$ norm corresponds to a more mixed state. Note that $H^{-1}$ norm can decrease in two ways: either by decreasing the amplitudes of $|\hat{\theta}_{\vec{k}}|$ or by advecting the flow such that the tracer concentration acquires more spectral mass in the higher wavenumbers and takes advantage of the $1/|\vec{k}|^2$ dependence. We use the $H^{-1}$ norm to define the (exponential) rate of mixing as
\begin{equation}
\label{eq:rate}
r(t) = -  \frac{\ddt{}\hmone{\theta}}{\hmone{\theta}}.
\end{equation}
The $L^{2}$ norm $\ltwo{\theta}$ and the $H^{1}$ norm $\hone{\theta}$ are also common measures of mixing and will be considered here as well.

We define the following ratio as a measure of the characteristic filamentation length scale:
\begin{equation}
\lambda(t)\equiv  2\pi \frac{\|\nabla^{-1}\theta(\,\cdot\,,t)\|_{L^{2}}}{\|\theta(\,\cdot\,,t)\|_{L^{2}}}.
\end{equation}
Note that $\lambda(t)$ returns the wavelength of the wavenumber $\vec{k}$ for a tracer concentration field composed only of the Fourier mode with wavenumber $\vec{k}$ (i.e. $\theta(\vec{x},t) = Re[ A e^{-i\vec{k}\cdot \vec{x}}]$ where $A$ is a complex constant). In general, $\lambda$ is the weighted root mean square wavelength with weights given by $|\theta_{\vec{k}}|/\ltwo{\theta}$.

For the enstrophy-bounded flow problem, we choose the same length scale $L$, the velocity scale $L\Gamma $, and  the time scale $1/\Gamma$. For the energy-bounded flow problem, we non-dimensionalize the system by choosing $L$ as the length scale, $U$ as the velocity scale, and $L/U$ as the time scale.  Both scalings produce the following form of the advection-diffusion equation,
\begin{equation}
\label{eq:nd_ade}
	\ppt{\theta}+\mathbf{u}\cdot \nabla \theta=\frac{1}{Pe} \lap\theta,
\end{equation}
where $Pe=  \frac{\Gamma L^2}{\kappa}$ for the enstrophy-constrained case and $Pe= \frac{UL}{\kappa}$ for the energy-constrained case.   The constraints on the flow become $\ltwo{\nabla\u} = 1$ or $\ltwo{\u} = 1$.

We will first review results for the $Pe = \infty$ enstrophy-constrained case. \cite{JFM2011} showed numerical evidence that the filament width $\lambda$ seemingly decreased continually by instanateous flow optimization. This type of optimization will be considered in this work as well. Self-similar flows can realize a decreasing $\lambda$ over time as shown analytically by \cite{Alberti2014a}.  It was rigorously proven indepedently by \cite{GI2014} and \cite{CS2013} that $\lambda$ decreased at most exponentially --- consistent with \cite{JFM2011}  and \cite{Alberti2014a}. The work of \cite{GI2014} relied on PDE regularization results of \cite{Crippa} while the approach of \cite{CS2013} used methods from optimal transportation theory \cite{villani2003topics}.

For $Pe = \infty$ energy-constrained problem, \cite{JMP2012} showed that a `checkerboard' flow can also provide continually decreasing length scales. Moreover, $\lambda$ decreased at a fast enough rate to approach zero in finite time in contrast to the enstrophy-constrained problem where $\lambda$ decreases to zero in infinite time.

The inclusion of diffusion ($Pe <\infty$) with the objective of optimal mixing has been explored.  \cite{DF2014} investigated optimal mixing and study the evolution of the mentioned measures ($H^{-1}, L^2,$ and $H^1$ norms) of mixing under the checkerboard flow introduced by \cite{JMP2012}. They show that the $H^{-1}$ and $L^2$ norms decrease monotonically under this flow while the $H^{1}$ increases until it reaches a peak and then decreases. This peak corresponds to a  time when the length scales developed are small enough for diffusion to effectively act on steep gradients. \cite{Miles2017a} explored this phenomena further in the context of a shell model, a reduced model using ordinary differential equations that mimic the spectral dynamics of the advection-diffusion equation. The authors concluded that shell-model mixing could not surpass length scales given by $\sqrt{\kappa/ \Gamma}$ for enstrophy-constrained flows and $\kappa/U$ for energy-constrained flows up to $O(1)$ constants. These length scales can be identified as a generalized Batchelor scale \cite{Batchelor1959a}, introduced in the context of turbulence theory.

In this work, we explore how diffusion impacts the evolution of the filament width $\lambda$ over time and mixing rates. We show numerical evidence that $\lambda$ appears to be limited by the Batchelor scale as seen in the shell model. Even when actively trying to choose the most optimal flow to minimize filimentation length. Thus, this may suggest that the Batchelor scale does not only limit turbulent flows but also all incompressible flows under the flow constraints considered here. Although these quantities have been known in the context of turbulence theory, the impact of these limitations on mixing rates has not been fully studied to our knowledge.

@@

The paper is organized as follows. We introduce the necessary theory regarding local-in-time optimization, a shell model, and $L^{\infty}$ flow constraints in section \ref{sec:theory}. Section \ref{sec:numerical_experiment} details the methodology and results of numerically implementing local-in-time flow optimization. Lastly, we finish with a discussion and conclusion in sections \ref{sec:discussion} and \ref{sec:conclusion} respectively.

\section{Theory}
\label{sec:theory}
\subsection{Local-in-time optimization}
We consider the local-in-time optimization strategy first introduced by \cite{JFM2011} in the diffusion-less case. We find that this strategy generalizes to the case with diffusion. The local-in-time optimal velocity fields maximize the instantaneous mixing rate by minimizing $\ddt{}\hmone{\theta}^2$ or equivalently minimizing $\ddt{\lambda^2}$ . The optimal velocity fields are given instantaneously by (in non-dimensional form)
%
\begin{equation}
\mathbf{u}= \frac{\mathds{P}(\theta \nabla \invlap\theta)}{\langle |\mathds{P}(\theta \nabla \invlap\theta)|^2\rangle^{1/2}}
\end{equation} 
%
for the energy constraint and by 
%
\begin{equation}
\mathbf{u}= \frac{-\invlap\mathds{P}(\theta \nabla \invlap\theta)}{\langle |\nabla^{-1}\mathds{P}(\theta \nabla \invlap\theta)|^2\rangle^{1/2}}
\end{equation}
%
for the enstrophy constraint where $\mathds{P}$ is the Leray divergence-free projector given by $\mathds{P}(\vec{v}) = \vec{v} - \nabla \Delta^{-1}(\nabla \cdot \vec{v})$. These flows will be studied numerically later and is the main focus of this paper.

\subsection{Shell model predictions of local-in-time optimization}

The shell model is a reduced model of the spectral dynamics present in the advection-diffusion equation. The model consists of a system of ordinary differential equations with nearest-neighboring coupling between `shells' in wavenumber space. \cite{Miles2017a} performed local-in-time mixing optimization in this model. The shell-model analysis predicts a limiting length scale given by the Batchelor scale, $\Lambda_{\Gamma} =\sqrt{\frac{\kappa}{\Gamma}}$  and its generalization $\Lambda_{U}= \frac{U}{\kappa} $. The non-dimensional versions are given by $\lambda_{\Gamma}= \frac{1}{\sqrt{Pe}}$ and $\lambda_{U} = \frac{1}{Pe}$.  Also from here forward, we will refer to the Batchelor scale to mean either $\lambda_{\Gamma}$ or its generalization $\lambda_{U}$.  The predicted long-term rates (after reaching the Batchelor scale) are given by $R_{\Gamma} =\kappa/\lambda_{\Gamma}^2 $  and  $R_{U}=\kappa/\lambda_{U}^2$  The non-dimensional versions are given by $r_{\Gamma} =1$ and $r_{U} = Pe $.

@@

Furthermore,
%
\numparts \begin{eqnarray}
\frac{d}{dt}\left(\frac{\|\nabla\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^2}\right) &= \frac{\|\theta\|_{L^{2}}^2\frac{d}{dt}\|\nabla\theta\|_{L^{2}}^2-\|\nabla\theta\|_{L^{2}}^2\frac{d}{dt}\|\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^4}\\
&= \frac{-2\int \partial_{i}u_{j}\partial_{i}\theta\partial_{j}\theta - \frac{2}{Pe} \|\Delta\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^2}+\frac{2}{Pe}\frac{\|\nabla\theta\|_{L^{2}}^4}{\|\theta\|_{L^{2}}^4} \\
&=-\frac{2}{Pe}\left(\frac{\|\Delta\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^2} - \frac{\|\nabla\theta\|_{L^{2}}^4}{\|\theta\|_{L^{2}}^4} \right) - 2\frac{\sint{\nabla\theta \cdot \nabla\vec{u} \cdot 
							\nabla\theta  }}{\|\theta\|_{L^{2}}^2} 
\\
&\leq 2 \frac{\hone{\theta}^2}{\ltwo{\theta}^2}
\end{eqnarray} \endnumparts
%
and using $\ddt{}\ltwo{\theta}^2 = -\frac{2}{Pe} \hone{\theta}^2$, it follows that
\begin{equation}
\ltwo{\theta}\geq  \ltwo{\theta_{0}}\exp\left[-\frac{1}{2Pe}\frac{\hone{\theta_{0}}^2}{\ltwo{\theta_{0}}^2}\left(e^{2 t} -1\right)\right].
\end{equation}
Using this with (\ref{eq:exponential_enstrophy}), we find that 
\begin{equation}
\hmone{\theta} \geq  \hmone{\theta_{0}} \exp\left[- t -\frac{1}{2 Pe}\frac{\hone{\theta_{0}}^2}{\ltwo{\theta_{0}}}\left(e^{2 t} -1\right)\right]
\end{equation}

\subsubsection{Results for $\linf{\u}= 1$}
Here we extend the results first introduced by \cite{Chi-Cheu1996}.  It is first useful to express the following quantities as:
%
\begin{eqnarray}
	 \hone{\theta}^2 &=& - 2\sint{\theta \lap \theta} \\
	 							&=& Pe \sint{\theta\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)} 
	 									-Pe \sint{\theta\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)} 
\end{eqnarray}
%
\begin{eqnarray}
	\ddt{}\ltwo{\theta}^2 &=& 2\sint{\theta\ppt{\theta}} \\
										 &=&\sint{\theta\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)} 
										 + \sint{\theta\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)} 
\end{eqnarray}
%
\begin{eqnarray}
	\ddt{}\hone{\theta}^2 &=& -2\sint{\ppt{\theta}\lap \theta} \\
	 									&=& Pe \sint{\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)^2} 
	 									-Pe \sint{\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)^2} .
\end{eqnarray}

We then compute the following derivative:
%
\begin{eqnarray*}
	\ddt{} \pbrac{ \frac{\hone{\theta}^2}{\ltwo{\theta}^2} } 
			&=& \frac{1}{\ltwo{\theta}^4}
			\sbrac{
				\ltwo{\theta}^2\ddt{}\hone{\theta}^2
				-\ddt{}\ltwo{\theta}^2\hone{\theta}^2			
			}\\
			&=& \frac{1}{\ltwo{\theta}^4}
			\sbrac{
				\ltwo{\theta}^2
				\pbrac{
					Pe \sint{\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)^2} 
 					-Pe\sint{\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)^2} 
				}
			}\\
		&-&\frac{1}{\ltwo{\theta}^4}
			\sbrac{
				Pe
				\pbrac{
					 \sint{\theta\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)} 
 				}^2		
 				-
 				Pe
 				\pbrac{
					 \sint{\theta\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)} 
 				}^2					
			}.
\end{eqnarray*}
%
Using H\"older's inequality and (\ref{eq:PDE_advection}), this simplifies to
%
\begin{eqnarray*}
	\ddt{} \pbrac{ \frac{\hone{\theta}^2}{\ltwo{\theta}^2} } 
			&\leq & \frac{Pe}{\ltwo{\theta}^2}
			\sbrac{
					 \sint{(\vec{u}\cdot \nabla \theta)^2} 
			}.
\end{eqnarray*}
By applying H\"older's inequality again, we have
\begin{equation}
\label{eq:k2growth_energy}
	\ddt{} 
		\pbrac{ 
			\frac{\hone{\theta}^2}{\ltwo{\theta}^2} 
		} 
		\leq  
		Pe
		\frac{\hone{\theta}^2}{\ltwo{\theta}^2} .
\end{equation}
%
Thus, we find that 
%
\begin{equation}
		\frac{\hone{\theta}}{\ltwo{\theta}} 
		\leq  
		\frac{\hone{\theta_0}}{\ltwo{\theta_0}}
		\exp{\pbrac{\frac{Pe}{2} t}}
\end{equation}
%
Using the inequality $\hone{\theta}\hmone{\theta}\geq \ltwo{\theta}^2$, we know that 
%
\begin{equation}
\label{eq:lambda_bound}
\lambda(t) \geq \frac{\ltwo{\theta_0}}{\hone{\theta_0}}\exp{\pbrac{-\frac{Pe}{2}t}}.
\end{equation}
%
Using (\ref{eq:k2growth_energy}) with  $\ddt{}\ltwo{\theta}^2 = -\frac{2}{Pe} \hone{\theta}^2$, 
we find that
%
\begin{equation}
\ltwo{\theta}\geq \ltwo{\theta_{0}}\exp\left[-\frac{1}{Pe^2}\frac{\hone{\theta_0}^2}{\ltwo{\theta_0}^2}\left(e^{Pe \, \, t}-1\right)\right]
\end{equation}
%
Using the above relation with  (\ref{eq:lambda_bound}), we arrive at
%
\begin{equation}
\hmone{\theta}\geq \frac{\ltwo{\theta_{0}}^2}{\hone{\theta_0}}\exp\left[-\frac{Pe}{2} \,\, t-\frac{1}{Pe^2}\frac{\hone{\theta_0}^2}{\ltwo{\theta_0}^2}\left(e^{Pe \,\, t}-1\right)\right]
\end{equation}

>>>>

\begin{abstract}
Incompressible flows can be effective mixers by appropriately advecting a passive tracer to produce small filamentation length scales. The Batchelor length scale \cite{Batchelor1959a} was proposed as the lower limit of length scales achievable in turbulent flows. Here, we provide numerical evidence that this limitation may be even more general, applying as well to all incompressible flows under mild physical constraints such as bounded enstrophy or energy. We consider local-in-time flow optimisation under these flow intensity constraints with the objective of maximising mixing rate performance. We observe that, for enstrophy-bounded optimal flows, the strength of diffusion has no impact on the long-term mixing rate performance. For energy-constrained optimal flows, however an increase in the strength of diffusion decreases the mixing rate. We provide analytical lower bounds on mixing rates and length scales achievable under related constraints (point-wise bounded speed and rate-of-strain) by extending the work of \cite{JFM2011} and \cite{Chi-Cheu1996}.

\end{abstract}
%
\pacs{47.85.lk, 47.85.L-, 42.10, 47.57.eb}
\ams{76R50, 37A25, 76F25, 76D55}
%
% Uncomment for keywords
\vspace{2pc}
\noindent{\it Keywords}: Mixing, incompressible flow, diffusion processes, flow control and optimisation, Batchelor scale

@@

Mixing a passive tracer concentration via an incompressible fluid flow is important to many areas of science and engineering. A question of interest across these domains is ``How can one stir to enhance the rate of mixing?'' We explore this question and study the impact of molecular diffusion. We address this topic by considering an incompressible flow with mild physical constraints in a periodic box with side length $L$ in $d$ dimensions. We consider a mean-zero tracer concentration field $\theta$ that evolves according to the advection-diffusion equation,
\begin{equation}
	\label{eq:PDE_advection}
	\ppt{\theta}+\mathbf{u}\cdot \nabla \theta=\kappa \lap\theta,
\end{equation}
with initial data $\theta(\mathbf{x},0)=\theta_{0}(\mathbf{x})$, where $\kappa$ is the molecular diffusion coefficient and $\mathbf{u}(\mathbf{x},t)$ is a incompressible ($\nabla\,\cdot\, \vec{u}=0$) flow field. The flow is constrained by enstrophy $\ltwo{\nabla\u} = \Gamma L^{d/2}$ or energy $\ltwo{\u} = UL^{d/2}$ where $\Gamma$ is the root mean square rate-of-strain and $U$ is the root mean square speed. We primarily use the $H^{-1}$ norm or mix-norm throughout to measure homogenisation,

\begin{equation}
\hmone{\theta}=\sqrt{\sint{ |\nabla^{-1} \theta( \vec{x},t)|^2}}=\sqrt{ \sum_{\vec{k}\neq \vec{0}} L^d \frac{|\hat{\theta}_{\vec{k}}(t)|^{2}}{|\vec{k}|^2}}
\end{equation}
where $\nabla^{-1}=\nabla \Delta^{-1}$, the operator $\Delta^{-1}$ acting on  a function $\rho$ returns the solution $\phi$ of the Poisson equation $ \Delta \phi = \rho $, and $\hat{\theta}_{\vec{k}}(t) =  \frac{1}{L^{d}}\sint{\theta(\vec{x},t)e^{-i\vec{k}\cdot\vec{x}}}$.  Lower values of the  $H^{-1}$ norm corresponds to a more mixed state. Note that $H^{-1}$ norm can decrease in two ways: either by decreasing the amplitudes of $|\hat{\theta}_{\vec{k}}|$ or by advecting the flow such that the tracer concentration acquires more spectral mass in the higher wave numbers and takes advantage of the $1/|\vec{k}|^2$ dependence. We use the $H^{-1}$ norm to define the (exponential) rate of mixing as
\begin{equation}
\label{eq:rate}
r(t) = -  \frac{\ddt{}\hmone{\theta}}{\hmone{\theta}}.
\end{equation}
The $L^{2}$ norm $\ltwo{\theta}$ and the $H^{1}$ norm $\hone{\theta}$ are also common measures of mixing and will be considered here as well.

We define the following ratio as a measure of the characteristic filamentation length scale:
\begin{equation}
\lambda(t)\equiv  2\pi \frac{\|\nabla^{-1}\theta(\,\cdot\,,t)\|_{L^{2}}}{\|\theta(\,\cdot\,,t)\|_{L^{2}}}.
\end{equation}
Note that $\lambda(t)$ returns the wavelength of the wave number $\vec{k}$ for a tracer concentration field composed only of the Fourier mode with wave number $\vec{k}$ (i.e. $\theta(\vec{x},t) = Re[ A e^{-i\vec{k}\cdot \vec{x}}]$ where $A$ is a complex constant). In general, $\lambda$ is the weighted root mean square wavelength with weights given by $|\theta_{\vec{k}}|/\ltwo{\theta}$.

For the enstrophy-bounded flow problem, we choose the same length scale $L$, the velocity scale $L\Gamma $, and  the time scale $1/\Gamma$. For the energy-bounded flow problem, we non-dimensionalise the system by choosing $L$ as the length scale, $U$ as the velocity scale, and $L/U$ as the time scale.  Both scalings produce the following form of the advection-diffusion equation,
\begin{equation}
\label{eq:nd_ade}
	\ppt{\theta}+\mathbf{u}\cdot \nabla \theta=\frac{1}{Pe} \lap\theta,
\end{equation}
where $Pe=  \frac{\Gamma L^2}{\kappa}$ for the enstrophy-constrained case and $Pe= \frac{UL}{\kappa}$ for the energy-constrained case.   The constraints on the flow become $\ltwo{\nabla\u} = 1$ or $\ltwo{\u} = 1$.

We will first review results for the $Pe = \infty$ enstrophy-constrained case. \cite{JFM2011} showed numerical evidence that the filament width $\lambda$ seemingly decreased continually by instantaneous flow optimisation. This type of optimisation will be considered in this work as well. Self-similar flows can realise a decreasing $\lambda$ over time as shown analytically by \cite{Alberti2014a}.  It was rigorously proven independently by \cite{GI2014} and \cite{CS2013} that $\lambda$ decreased at most exponentially --- consistent with \cite{JFM2011}  and \cite{Alberti2014a}. The work of \cite{GI2014} relied on PDE regularisation results of \cite{Crippa} while the approach of \cite{CS2013} used methods from optimal transportation theory \cite{villani2003topics}.

For $Pe = \infty$ energy-constrained problem, \cite{JMP2012} showed that a `chequerboard' flow can also provide continually decreasing length scales. Moreover, $\lambda$ decreased at a fast enough rate to approach zero in finite time in contrast to the enstrophy-constrained problem where $\lambda$ decreases to zero in infinite time.

The inclusion of diffusion ($Pe <\infty$) with the objective of optimal mixing has been explored.  \cite{DF2014} investigated optimal mixing and study the evolution of the mentioned measures ($H^{-1}, L^2,$ and $H^1$ norms) of mixing under the chequerboard flow introduced by \cite{JMP2012}. They show that the $H^{-1}$ and $L^2$ norms decrease monotonically under this flow while the $H^{1}$ increases until it reaches a peak and then decreases. This peak corresponds to a  time when the length scales developed are small enough for diffusion to effectively act on steep gradients. \cite{Miles2017a} explored this phenomena further in the context of a shell model, a reduced model using ordinary differential equations that mimic the spectral dynamics of the advection-diffusion equation. The authors concluded that shell-model mixing could not surpass length scales given by $\sqrt{\kappa/ \Gamma}$ for enstrophy-constrained flows and $\kappa/U$ for energy-constrained flows up to $O(1)$ constants. These length scales can be identified as a generalised Batchelor scale \cite{Batchelor1959a}, introduced in the context of turbulence theory.

In this work, we explore how diffusion impacts the evolution of the filament width $\lambda$ over time and mixing rates. We show numerical evidence that $\lambda$ appears to be limited by the Batchelor scale as seen in the shell model. Even when actively trying to choose the most optimal flow to minimise filamentation length. Thus, this may suggest that the Batchelor scale does not only limit turbulent flows but also all incompressible flows under the flow constraints considered here. Although these quantities have been known in the context of turbulence theory, the impact of these limitations on mixing rates has not been fully studied to our knowledge.

@@

The paper is organised as follows. We introduce the necessary theory regarding local-in-time optimisation, a shell model, and $L^{\infty}$ flow constraints in section \ref{sec:theory}. Section \ref{sec:numerical_experiment} details the methodology and results of numerically implementing local-in-time flow optimisation. Lastly, we finish with a discussion and conclusion in sections \ref{sec:discussion} and \ref{sec:conclusion} respectively.

\section{Theory}
\label{sec:theory}
\subsection{Local-in-time optimisation}
We consider the local-in-time optimisation strategy first introduced by \cite{JFM2011} in the diffusion-less case. We find that this strategy generalises to the case with diffusion. The local-in-time optimal velocity fields maximise the instantaneous mixing rate by minimising $\ddt{}\hmone{\theta}^2$ or equivalently minimising $\ddt{\lambda^2}$ . The optimal velocity fields are given instantaneously by (in non-dimensional form)
%
\begin{equation}
\mathbf{u}= \frac{\mathds{P}(\theta \nabla \invlap\theta)}{\langle |\mathds{P}(\theta \nabla \invlap\theta)|^2\rangle^{1/2}}
\end{equation} 
%
for the energy constraint and by 
%
\begin{equation}
\mathbf{u}= \frac{-\invlap\mathds{P}(\theta \nabla \invlap\theta)}{\langle |\nabla^{-1}\mathds{P}(\theta \nabla \invlap\theta)|^2\rangle^{1/2}}
\end{equation}
%
for the enstrophy constraint where $\mathds{P}$ is the Leray divergence-free projector given by $\mathds{P}(\vec{v}) = \vec{v} - \nabla \Delta^{-1}(\nabla \cdot \vec{v})$. These flows will be studied numerically later and is the main focus of this paper.

\subsection{Shell model predictions of local-in-time optimisation}

The shell model is a reduced model of the spectral dynamics present in the advection-diffusion equation. The model consists of a system of ordinary differential equations with nearest-neighbour coupling between `shells' in wave number space. \cite{Miles2017a} performed local-in-time mixing optimisation in this model. The shell-model analysis predicts a limiting length scale given by the Batchelor scale, $\Lambda_{\Gamma} =\sqrt{\frac{\kappa}{\Gamma}}$  and its generalisation $\Lambda_{U}= \frac{U}{\kappa} $. The non-dimensional versions are given by $\lambda_{\Gamma}= \frac{1}{\sqrt{Pe}}$ and $\lambda_{U} = \frac{1}{Pe}$.  Also from here forward, we will refer to the Batchelor scale to mean either $\lambda_{\Gamma}$ or its generalisation $\lambda_{U}$.  The predicted long-term rates (after reaching the Batchelor scale) are given by $R_{\Gamma} =\kappa/\lambda_{\Gamma}^2 $  and  $R_{U}=\kappa/\lambda_{U}^2$  The non-dimensional versions are given by $r_{\Gamma} =1$ and $r_{U} = Pe $.

@@

Furthermore,
%
\numparts \begin{eqnarray}
\frac{d}{dt}\left(\frac{\|\nabla\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^2}\right) &= \frac{\|\theta\|_{L^{2}}^2\frac{d}{dt}\|\nabla\theta\|_{L^{2}}^2-\|\nabla\theta\|_{L^{2}}^2\frac{d}{dt}\|\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^4}\\
&= \frac{-2\int \partial_{i}u_{j}\partial_{i}\theta\partial_{j}\theta - \frac{2}{Pe} \|\Delta\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^2}+\frac{2}{Pe}\frac{\|\nabla\theta\|_{L^{2}}^4}{\|\theta\|_{L^{2}}^4} \\
&=-\frac{2}{Pe}\left(\frac{\|\Delta\theta\|_{L^{2}}^2}{\|\theta\|_{L^{2}}^2} - \frac{\|\nabla\theta\|_{L^{2}}^4}{\|\theta\|_{L^{2}}^4} \right) - 2\frac{\sint{\nabla\theta \cdot \nabla\vec{u} \cdot 
							\nabla\theta  }}{\|\theta\|_{L^{2}}^2} 
\\
&\leq 2 \frac{\hone{\theta}^2}{\ltwo{\theta}^2}
\end{eqnarray} \endnumparts
%
and using $\ddt{}\ltwo{\theta}^2 = -\frac{2}{Pe} \hone{\theta}^2$, it follows that
\begin{equation}
\ltwo{\theta}\geq  \ltwo{\theta_{0}}\exp\left[-\frac{1}{2Pe}\frac{\hone{\theta_{0}}^2}{\ltwo{\theta_{0}}^2}\left(e^{2 t} -1\right)\right].
\end{equation}
Using this with (\ref{eq:exponential_enstrophy}), we deduce the lower bound
\begin{equation}
\hmone{\theta} \geq  \hmone{\theta_{0}} \exp\left[- t -\frac{1}{2 Pe}\frac{\hone{\theta_{0}}^2}{\ltwo{\theta_{0}}}\left(e^{2 t} -1\right)\right].
\end{equation}

\subsubsection{Results for $\linf{\u}= 1$}
Here we extend the result of \cite{Chi-Cheu1996} to show that the presence of diffusion rules out perfect mixing in finite time for bounded velocity flows.  First note that
%
\begin{eqnarray}
	 \hone{\theta}^2 &=& - 2\sint{\theta \lap \theta} \\
	 							&=& Pe \sint{\theta\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)} 
	 									-Pe \sint{\theta\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)} 
\end{eqnarray},
%
\begin{eqnarray}
	\ddt{}\ltwo{\theta}^2 &=& 2\sint{\theta\ppt{\theta}} \\
										 &=&\sint{\theta\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)} 
										 + \sint{\theta\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)} 
\end{eqnarray},
%
and
%
\begin{eqnarray}
	\ddt{}\hone{\theta}^2 &=& -2\sint{\ppt{\theta}\lap \theta} \\
	 									&=& Pe \sint{\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)^2} 
	 									-Pe \sint{\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)^2} .
\end{eqnarray}

Then compute:
%
\begin{eqnarray*}
	\ddt{} \pbrac{ \frac{\hone{\theta}^2}{\ltwo{\theta}^2} } 
			&=& \frac{1}{\ltwo{\theta}^4}
			\sbrac{
				\ltwo{\theta}^2\ddt{}\hone{\theta}^2
				-\ddt{}\ltwo{\theta}^2\hone{\theta}^2			
			}\\
			&=& \frac{1}{\ltwo{\theta}^4}
			\sbrac{
				\ltwo{\theta}^2
				\pbrac{
					Pe \sint{\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)^2} 
 					-Pe\sint{\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)^2} 
				}
			}\\
		&-&\frac{1}{\ltwo{\theta}^4}
			\sbrac{
				Pe
				\pbrac{
					 \sint{\theta\left(\ppt{\theta}
	 									-\frac{1}{Pe}\lap \theta\right)} 
 				}^2		
 				-
 				Pe
 				\pbrac{
					 \sint{\theta\left(\ppt{\theta}
	 									+\frac{1}{Pe}\lap \theta\right)} 
 				}^2					
			}.
\end{eqnarray*}
%
Using H\"older's inequality and (\ref{eq:PDE_advection}), this simplifies to
%
\begin{eqnarray*}
	\ddt{} \pbrac{ \frac{\hone{\theta}^2}{\ltwo{\theta}^2} } 
			&\leq & \frac{Pe}{\ltwo{\theta}^2}
			\sbrac{
					 \sint{(\vec{u}\cdot \nabla \theta)^2} 
			}.
\end{eqnarray*}
Again applying H\"older's inequality we have
\begin{equation}
\label{eq:k2growth_energy}
	\ddt{} 
		\pbrac{ 
			\frac{\hone{\theta}^2}{\ltwo{\theta}^2} 
		} 
		\leq  
		Pe
		\frac{\hone{\theta}^2}{\ltwo{\theta}^2} 
\end{equation}
%
and thus 
%
\begin{equation}
		\frac{\hone{\theta}}{\ltwo{\theta}} 
		\leq  
		\frac{\hone{\theta_0}}{\ltwo{\theta_0}}
		\exp{\pbrac{\frac{Pe}{2} t}}.
\end{equation}
%
The inequality $\hone{\theta}\hmone{\theta}\geq \ltwo{\theta}^2$ then ensures that
%
\begin{equation}
\label{eq:lambda_bound}
\lambda(t) \geq \frac{\ltwo{\theta_0}}{\hone{\theta_0}}\exp{\pbrac{-\frac{Pe}{2}t}}.
\end{equation}
%
Using (\ref{eq:k2growth_energy}) together with  $\ddt{}\ltwo{\theta}^2 = -\frac{2}{Pe} \hone{\theta}^2$ we observe that
%
\begin{equation}
\ltwo{\theta}\geq \ltwo{\theta_{0}}\exp\left[-\frac{1}{Pe^2}\frac{\hone{\theta_0}^2}{\ltwo{\theta_0}^2}\left(e^{Pe \, \, t}-1\right)\right]
\end{equation}
%
and this combined with  (\ref{eq:lambda_bound}) implies
%
\begin{equation}
\hmone{\theta}\geq \frac{\ltwo{\theta_{0}}^2}{\hone{\theta_0}}\exp\left[-\frac{Pe}{2} \,\, t-\frac{1}{Pe^2}\frac{\hone{\theta_0}^2}{\ltwo{\theta_0}^2}\left(e^{Pe \,\, t}-1\right)\right].
\end{equation}
